---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# qtl2hotspots





<!-- badges: start -->
[![Travis-CI Build Status](https://travis-ci.org/fboehm/qtl2hotspots.svg?branch=master)](https://travis-ci.org/fboehm/qtl2hotspots)

<!-- badges: end -->

The goal of `qtl2hotspots` is to analyze the Keller et al. (2018) expression trait hotspots with both mediation methods and pleiotropy tests.

## Installation

Install from Dockerhub with this command:

```{bash, docker_pull, eval = FALSE}
docker pull fjboehm/qtl2hotspots
```

Alternatively, if you wish to install the package directly from Github, use `devtools`.

```{r, devtools_stuff, eval = FALSE}
install.packages("devtools")
devtools::install_github("fboehm/qtl2hotspots")
```

## Keller, et al. (2018)'s hotspots table

Here is Keller's Table S1 (from the Supplemental Documents).

| Chr | position (Mb) |  number of transcripts  | candidate mediator  | number of genes with LOD difference > 1.5  |     
| --- | -----------   |  ---------------------  |  ------------------ |  ----------------------------------------- |   
| 2    | 165.5        |   147                   |  Hnf4a              |  88                                        | 
| 5    | 146          |   182                   |  Pdx1               |  77                                        |
| 7    | 46           |   123                   |  Fam83e             |  96                                        |
| 11   | 71           |   126                   |  Sat2               |  115                                       |
| 13   | 112.5        |   104                   |  Il6st              |  82                                        |

Table: From [Keller et al. 2018 GENETICS](https://www.genetics.org/content/209/1/335), [Supplementary table 1](https://figshare.com/articles/Supplemental_Material_for_Attie_et_al_2018_in_review_/5977459)




## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
map <- readRDS(here::here("analysis", "data", "derived_data", "map.rds"))
hot_chr <- c(2, 5, 7, 11, 13)
start_index <- c(3742, 3498, 927, 1868, 2814) # from s1 values in condor submit files
scan_length <- c(247, 269, 242, 217, 226) # in number of markers, from condor submit files
stop_index <- start_index + scan_length - 1 # index of the last marker in the interval
start_position <- c(map$`2`[start_index[1]], 
                    map$`5`[start_index[2]], 
                    map$`7`[start_index[3]],
                    map$`11`[start_index[4]],
                    map$`13`[start_index[5]]
)
stop_position <- c(map$`2`[stop_index[1]], 
                    map$`5`[stop_index[2]], 
                    map$`7`[stop_index[3]],
                    map$`11`[stop_index[4]],
                    map$`13`[stop_index[5]]
)
tibble::tibble(hot_chr, start_index, scan_length, stop_index, start_position, stop_position)
```

## Build the manuscript `paper.Rnw`

To render the manuscript `paper.Rnw`, into pdf, use a shell to enter the following text from the `qtl2hotspots` directory:

```{bash, docker_build, eval = FALSE}
docker build .
```

If you're not using Docker, and have cloned the repository from Github, you can use the Makefile to render `paper.Rnw`. Just type `make` in the shell (from within the `qtl2hotspots` directory):

```{bash, make, eval = FALSE}
make
```




## Code of Conduct

Please note that the `qtl2hotspots` project is released with a
[Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md).
By contributing to this project, you agree to abide by its terms.

