---
title: "Expression QTL hotspot dissection with mediation analysis and pleiotropy testing"
author: "Frederick J. Boehm, Brian S. Yandell, and Karl W. Broman"
date: "7/27/2019"
bibliography: research.bib
numberSections: true
linkReferences: true
link-citations: true
output: 
  bookdown::html_document2:
    pandoc_args: 
      - --filter
      - /usr/local/Cellar/pandoc-crossref/0.3.4.1/bin/pandoc-crossref
---

```{r child="intro.Rmd"}
```


```{r load-broman}
library(broman)
library(here)
```


```{r child="methods.Rmd"}
```






# Results

We generated scatter plots of pleiotropy test statistics vs. mediation LOD difference values for each of the five hotspots (Figures).
Each point represents a pairing of a local trait with a nonlocal trait.
The mediation LOD difference value is that observed when the local trait is considered
as a candidate mediator of the QTL-nonlocal trait association.

In all five scatter plots, we observe few
points in the upper right quadrant of the figure.
That is, few pairs have high pleiotropy test statistics and
high mediation LOD difference values.

```{r setup}
library(qtl2hotspots)
pleio_threshold <- 4
ult <- 15 # univariate lod threshold
```

```{r load_tidyverse}
library(tidyverse)
options(tibble.print_max = 50, tibble.print_min = 20)
```


```{r load_heatmaply}
library(heatmaply)
```


```{r setup_continued-2}
hot_chr <- 2
hot_mid <- 165.5
keller_mediator <- "Hnf4a"
```

```{r child="hotspot-one-chromosome.Rmd"}
```



```{r, tmem189-scatter, fig.cap="Colors correspond to mediation by Tmem189"}
bar <- foo %>%
  filter(local_symbol == "Tmem189")

p <- tibble(nonlocal_symbol = rownames(pm[indic_nonlocal, indic_nonlocal])[ord_nonlocal], index = 1:length(nonlocal_symbol)) %>%
  inner_join(bar, by = "nonlocal_symbol") %>% 
  left_join(nonlocal_annot, by = c("nonlocal_symbol" = "symbol")) %>%
  ggplot() + geom_point(aes(y = pos, x = index, color = lod_diff > 1.5, nonlocal = nonlocal_symbol, lod_diff = lod_diff)) 

ggplotly(p)
```



## Chr 2 hotspot results






```{r setup_continued-5}
hot_chr <- 5
hot_mid <- 146
keller_mediator <- "Pdx1"
```

```{r child="hotspot-one-chromosome.Rmd", eval = FALSE}
```





```{r setup_continued-7}
hot_chr <- 7
hot_mid <- 46
keller_mediator <- "Fam83e"
```

```{r child="hotspot-one-chromosome.Rmd", eval = FALSE}
```



```{r setup_continued-11}
hot_chr <- 11
hot_mid <- 71
keller_mediator <- "Sat2"
```

```{r child="hotspot-one-chromosome.Rmd", eval = FALSE}
```



```{r setup_continued-13}
hot_chr <- 13
hot_mid <- 112.5
keller_mediator <- "Il6st"
```

```{r child="hotspot-one-chromosome.Rmd", eval = FALSE}
```











# Discussion

Pleiotropy tests complement mediation analyses in the following manner. When considering a
collection of potential mediators for a nonlocal trait, mediation analyses clearly identify the
mediator when there is, in fact, a mediator among the candidates. However, when there is no
mediator, mediation analyses provide little information about the genetics of the complex trait.
Pleiotropy testing is particularly useful in this setting, since a pleiotropy test always gives
an answer about the number of underlying QTL. A drawback of pleiotropy testing is that the scientific question that it addresses - do two traits share a single QTL - informs, but doesn't fully resolve, the genetics of the nonlocal complex traits under study.









# Colophon

This report was generated on ` r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE, echo = TRUE, tidy.opts=list(width.cutoff=50), size = "tiny", results = "markup"}
# which R packages and versions?
devtools::session_info()
```


The current Git commit details are:

```{r git2r, cache = FALSE, echo = TRUE}
# what commit is this file at?
git2r::repository(here::here())
```




