---
title: "Checking heatmaps and dendrograms"
author: "Frederick Boehm"
date: "7/14/2019"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(hotspots)
hot_chr <- 11
hot_mid <- 71
keller_mediator <- "Sat2"
```

```{r more_setup}
hot_lower <- hot_mid - 2
hot_upper <- hot_mid + 2
fp <- paste0("../figures/Chr", hot_chr, "-")
knitr::opts_chunk$set(fig.path = fp)
```

The questions that I want to address today deal with heatmaps and dendrograms. Basically, I want to ensure that I'm correctly sorting & ordering a data matrix before passing it to a heatmap function.

I need to check a few things. First, what is the ordering of rows of a matrix when I create a dendrogram? What is the ordering of rows when I create a heatmap (without attempting any manual reordering)?

Can I order the data matrix according to the dendrogram so that, when calling the heatmap function *while preventing the automated reordering of rows*, do I get the same ordering as in the dendrogram?

Other questions that I want to consider include the *method* for dendrogram creation. I also want to think about coloring schemes in heatmaps.

## Reading Chr 11 intermediate result files 

Let's examine the Chr 11 files for pleiotropy results.

```{r readRDS}
pleio_mixed_list <- readRDS("_cache/Chr11_pleio.rds")
pleio_local_list <- readRDS("_cache/Chr11_pleio_local.rds")
pleio_nonlocal_list <- readRDS("_cache/Chr11_pleio_nonlocal.rds")
med_list <- readRDS("_cache/Chr11_mediations.rds")
```

```{r, load_attie}
load("~/Box Sync/attie/keller2018-chr2-hotspot-chtc/data-to-ignore/Attie_DO378_eQTL_viewer_v1.Rdata")
library(tidyverse)
options(tibble.print_max = 50, tibble.print_min = 20)
```


```{r make_local_annot}
local_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>%
  mutate(local_tx = chr == hot_chr) %>%
  filter(local_tx)
```


```{r make_nonlocal_annot}
nonlocal_annot <- dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == hot_chr, pos <= hot_upper, pos >= hot_lower) %>%
  left_join(y = dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(!duplicated(annot.id)) %>%
  filter(lod >= 7.18) %>% #genomewide threshold
  mutate(local_tx = chr == hot_chr) %>%
  filter(!local_tx) %>% 
  filter(!is.na(hotspot))
```


```{r, make_local_annot2}
local_annot2 <- local_annot %>%
  filter(middle <= hot_upper + 5, middle >= hot_lower - 5) %>%
  arrange(desc(lod))
```

```{r subset_expr_matrix}
hot_local <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% local_annot2$annot.id]
hot_nonlocal <- dataset.islet.rnaseq$expr[, colnames(dataset.islet.rnaseq$expr) %in% nonlocal_annot$annot.id]
nnonlocal <- ncol(hot_nonlocal)
nlocal <- ncol(hot_local)
```




```{r tabulate, message=FALSE, warning = FALSE}
med_tib <- med_list %>%
  purrr::map(bind_rows) %>%
  bind_rows() %>%
  as_tibble() %>%
  rename(lod_without_med = lod_no_med, lod_with_med = lod_med) %>%
  mutate(lod_diff = lod_without_med - lod_with_med, lod_diff_proportion = lod_diff / lod_without_med) %>%
  mutate(local_gene_id = rep(colnames(hot_local), each = nnonlocal), nonlocal_gene_id = rep(colnames(hot_nonlocal), times = nlocal)) %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  left_join(nonlocal_annot, by = c("nonlocal_gene_id" = "annot.id")) %>%
  dplyr::select(symbol.x, symbol.y, lod_without_med, lod_with_med, lod_diff, lod_diff_proportion, local_gene_id, nonlocal_gene_id) %>%
  rename(local_symbol = symbol.x, nonlocal_symbol = symbol.y)
```






```{r make_pleio_tib}
hot_annot <- local_annot2 %>%
  bind_rows(nonlocal_annot) %>%
  select(annot.id, symbol)
(pleio_tib <- pleio_mixed_list %>%
  purrr::map(.f = function(x) rename(x, gene1_id = local_gene_id, gene2_id = nonlocal_gene_id)) %>%
  c(pleio_local_list, pleio_nonlocal_list) %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(hot_annot, by = c("gene1_id" = "annot.id")) %>%
  rename(gene1_symbol = symbol) %>%
  left_join(hot_annot, by = c("gene2_id" = "annot.id")) %>%
  rename(gene2_symbol = symbol) %>%
  select(gene1_symbol, gene2_symbol, pleiotropy_lod)
)  
```

```{r hier_clust}
pm <- pleio_tib %>%
  tibble_to_matrix(symmetric = TRUE)
hc <- pm %>% 
  dist() %>%
  hclust()
hc %>%
  plot(cex = 0.2)
```

```{r}
hc$labels == rownames(pm)[hc$order]
```

```{r}
tibble(hc$labels, rownames(pm)[hc$order], hc$labels[hc$order], rownames(pm))
```

The tibble above shows the needed result. `hc$labels` is, by default definition, equal to `rownames(pm)`, while `rownames(pm)[hc$order]` gives the sequence of symbols that align with the leaves of the dendrogram.

```{r load_heatmaply}
library(heatmaply)
```

The vignette for `heatmaply` is quite helpful and includes discussion of its methods for ordering rows.

```{r heat-pleio}
heatmaply(pm, symm = TRUE)
```

```{r heat1}
heatmaply(percentize(pm), cexRow = 0.2, cexCol = 0.2, seriate = "OLO", symm = TRUE)
```

```{r heat2}
heatmaply(percentize(pm), cexRow = 0.2, cexCol = 0.2, seriate = "mean", symm = TRUE)
```

```{r heat3}
heatmaply(normalize(pm), 
          cexRow = 0.2, 
          cexCol = 0.2, 
          seriate = "OLO", 
          symm = TRUE, dendrogram = "both"
          )
```

## Heatmaps for matrix containing only nonlocal traits

```{r heat4}
pm_nonlocal <- pleio_tib %>%
  filter(gene1_symbol %in% nonlocal_annot$symbol, gene2_symbol %in% nonlocal_annot$symbol) %>%
  tibble_to_matrix(symmetric = TRUE)
pm_nonlocal %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)

```

```{r heat5}
(pm_nonlocal > 3) %>%
  apply(FUN = as.integer, MARGIN = 2) %>%
  heatmaply(symm = TRUE, cexRow = 0.3, cexCol = 0.3)
```


In the matrix of nonlocal - nonlocal pairs of pleiotropy test statistics, the fact that the maximum is only `r round(max(pm_nonlocal, na.rm = TRUE))` tells us that the heatmap of binary values is not going to be informative, since all or nearly all entries are less than any reasonable threshold.



